Exploring the Impact of Population Density on Crime Distribution in Toronto Neighborhoods (2014–2023)

### 2024-09-26 paper/paper.qmd
Update LLM usage
Delete unused file.

# Introduction
Urban crime trends have long been studied due to their impact on social stability and policy-making. In Toronto, crime rates vary significantly across neighborhoods, often influenced by factors such as population density and economic disparity. This paper focuses on analyzing the trends in crime frequency across different types of crimes, using data from 2014 to 2023. The objective is to determine how crime patterns have changed over time and which neighborhoods are most affected. The result shows a icnreasing trend in crime rates. All the analysis in the paper is conducted in @citeR. (Need to write more!)

@@

## Geographic Distribution of Crime
Based on @fig-Population, @fig-percrime represents a higher concentration of crimes in areas with higher population densities, particularly in economically disadvantaged neighborhoods. Those gray areas mean no crime was reported. From the map showing **Homicide**, it This finding shows that population size is a strong predictor of crime, with more densely populated areas experiencing higher crime rates. In Toronto, eastern neighborhoods report higher assault and robbery rates than western, more affluent areas.

>>>>

# Introduction
Urban crime trends have long been studied due to their impact on social stability and policy-making. In Toronto, crime rates vary significantly across neighborhoods, often influenced by factors such as population density and economic disparity. This paper focuses on analyzing the trends in crime frequency across different types of crimes, using data from 2014 to 2023. The objective is to determine how crime patterns have changed over time and which neighborhoods are most affected. The result shows a increasing trend in crime rates. All the analysis in the paper is conducted in @citeR, @openai2023chatgpt is used in this research paper to help simulate data. (Need to write more!)

@@

## Geographic Distribution of Crime
Based on @fig-Population, @fig-percrime represents a higher concentration of crimes in areas with higher population densities, particularly in economically disadvantaged neighborhoods. Those gray areas mean no crime was reported. From the map showing **Homicide**, it This finding shows that population size is a strong predictor of crime, with more densely populated areas experiencing higher crime rates. In Toronto, eastern neighborhoods report higher assault and robbery rates than western, more affluent areas.
```{r,fig.width=10, fig.height=8}
#| label: fig-crimeandpopulation
#| fig-cap: 2023 Crime Trend with Population
#| echo: false
#| warning: false
#| message: false
crime_long_1 <- read_rds(here::here("data/analysis_data/cleaning_crime_count.rds"))

ggplot(crime_long_1, aes(x = Crime_Count, y = POPULATION_2023)) +
  geom_point(aes(color = Crime_Type), size = 2, alpha = 0.8) +  # Adjust point size and transparency for better visibility
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1) +  # Keep linear trendline with original method
  scale_color_manual(values = c("lightpink", "red", "grey", "blue", "orange", "skyblue", "brown", "yellow")) +  # Customize colors for crime types
  scale_x_continuous(limits = c(0, 200), breaks = seq(0, 200, by = 50)) +  # Set horizontal axis limits from 0 to 200 with breaks
  facet_wrap(~ Crime_Type) +  # Split the scatterplot by crime type
  labs(title = "Crime Count vs. Population by Neighborhoods",
       subtitle = "Visualized for Multiple Crime Types",
       x = "Crime Count",  
       y = "Population",
       color = "Crime Type") +
  theme_minimal(base_size = 15) +  # Increase base size for text readability
  theme(axis.title = element_text(size = 14),  
        panel.grid.major = element_line(color = "gray90"),  # Add subtle grid lines for clarity
        panel.grid.minor = element_blank(),   
        strip.text = element_text(size = 16, face = "bold"),  # Emphasize facet labels
        legend.position = "bottom",  # Move legend to the bottom for a cleaner layout
        legend.title = element_text(face = "bold"),
        plot.title = element_text(face = "bold", hjust = 0.5),  # Center align the title
        plot.subtitle = element_text(hjust = 0.5))  # Center align the subtitle
```
